---
title: "UE20CS312 - Data Analytics - Worksheet 1a - Part 2 - EDA with R | ANOVA"
subtitle: "PES University"
author: "'SUNDEEP A, Dept. of CSE - PES1UG20CS445'"
date: "2022-08-26"
output: pdf_document
urlcolor: blue
editor_options:
markdown:
wrap: 72
---

# Part I. Exploratory Data Analysis with R

## Book Club Marketing Dataset

## Problems

### Problem 1 (1 point)

```{r}
library(tidyverse)
cbc_df <- read_csv("C:\\Users\\HP\\Desktop\\PESU\\SEM - 5\\PES1UG20CS445\\DA\\1_b\\CharlesBookClubDataset.csv")
head(cbc_df)
summary(cbc_df[c("M","F","R","FirstPurch","ChildBks","YouthBks","CookBks","DoItYBks","RefBks","ArtBks","GeogBks","ItalCook","ItalAtlas","ItalArt","Florence","Related Purchase","Yes_Florence","No_Florence")])
sprintf("The standard Deviation of M data : %f",sd(cbc_df$M,na.rm = T))
sprintf("The standard Deviation of F data : %f",sd(cbc_df$F,na.rm = T))
sprintf("The standard Deviation of R data : %f",sd(cbc_df$R,na.rm = T))
sprintf("The standard Deviation of FirstPurch data : %f",sd(cbc_df$FirstPurch))
sprintf("The standard Deviation of ChildBks data : %f",sd(cbc_df$ChildBks))
sprintf("The standard Deviation of YouthBks data : %f",sd(cbc_df$YouthBks))
sprintf("The standard Deviation of CookBks data : %f",sd(cbc_df$CookBks))
sprintf("The standard Deviation of DoItYBks data : %f",sd(cbc_df$DoItYBks))
sprintf("The standard Deviation of RefBks data : %f",sd(cbc_df$RefBks))
sprintf("The standard Deviation of ArtBks data : %f",sd(cbc_df$ArtBks))
sprintf("The standard Deviation of GeogBks data : %f",sd(cbc_df$GeogBks))
sprintf("The standard Deviation of ItalCook data : %f",sd(cbc_df$ItalCook))
sprintf("The standard Deviation of ItalAtlas data : %f",sd(cbc_df$ItalAtlas))
sprintf("The standard Deviation of ItalArt data : %f",sd(cbc_df$ItalArt))
sprintf("The standard Deviation of Florence data : %f",sd(cbc_df$Florence))
sprintf("The standard Deviation of Related Purchase data : %f",sd(cbc_df$`Related Purchase`))
sprintf("The standard Deviation of Yes_Florence data : %f",sd(cbc_df$Yes_Florence))
sprintf("The standard Deviation of NO_Florence data : %f",sd(cbc_df$No_Florence))
print("Count per each column : ")
print(colSums(!is.na(cbc_df)))
print("Number of missing Values per column :")
colSums(is.na(cbc_df))
```

### Problem 2 (2 points)

```{r}
hist(cbc_df$R)
cbc_df$R[is.na(cbc_df$R)] <- median(cbc_df$R, na.rm = T)
hist(cbc_df$R)
hist(cbc_df$F)
cbc_df$F[is.na(cbc_df$F)] <- median(cbc_df$F, na.rm = T)
hist(cbc_df$F)
hist(cbc_df$M)
cbc_df$M[is.na(cbc_df$M)] <- mean(cbc_df$M, na.rm = T)
hist(cbc_df$M)
print("So we can observe that there are no missing values")
colSums(is.na(cbc_df))
```

## Analysis:

Since the Histogram of R dataset is Right Skewed, So we can replace the missing values with the **median** of the R data

Similarly the Histogram of F dataset is Right Skewed, So we can replace the missing values with the **median** of the F data

Finally the Histogram of the M dataset is normally distributed,So we can replace the missing Values with **mean** of the M data

## Problem 3 (2 points)

```{r}
cbc_df$Mcode <- cut(cbc_df$M, breaks = 5)
cbc_df$Rcode <- cut(cbc_df$R, breaks = 5)
cbc_df$Fcode <- cut(cbc_df$F, breaks = 5)

summary(cbc_df[c("Mcode","Rcode","Fcode")])

```

### Problem 4

#### 4.1 Bar Graphs (1 point)

```{r}
hist(cbc_df$FirstPurch)
numeric_Mcode=as.numeric(cbc_df$Mcode)
numeric_Rcode=as.numeric(cbc_df$Rcode)
numeric_Fcode=as.numeric(cbc_df$Fcode)
hist(numeric_Mcode)
hist(numeric_Fcode)
hist(numeric_Rcode)
```

#### 4.2 Box Plot (1 point)

```{r}
cbc_df$Florence=factor(cbc_df$Florence, labels=c("TRUE","FALSE"))
F_R_T=cbc_df$R[which(cbc_df$Florence=="TRUE")]
F_R_F=cbc_df$R[which(cbc_df$Florence=="FALSE")]
boxplot(F_R_T,F_R_F,horizontal = TRUE,xlab="R",ylab="Florence",names = c("TRUE","FALSE"))

F_M_T=cbc_df$M[which(cbc_df$Florence=="TRUE")]
F_M_F=cbc_df$M[which(cbc_df$Florence=="FALSE")]
boxplot(F_M_T,F_M_F,horizontal = TRUE,xlab="M",ylab="Florence",names = c("TRUE","FALSE"))

F_First_T=cbc_df$FirstPurch[which(cbc_df$Florence=="TRUE")]
F_First_F=cbc_df$FirstPurch[which(cbc_df$Florence=="FALSE")]
boxplot(F_First_T,F_First_F,horizontal = TRUE,xlab="FirstPurch",ylab="Florence",names = c("TRUE","FALSE"))
```

#### 4.3 Density Plot (1 point)

```{r}
plot(density(cbc_df$R),main="Recency")
plot(density(cbc_df$F),main="Frequency")
plot(density(cbc_df$M),main="Monetary")
plot(density(cbc_df$FirstPurch),main="FirstPurch")

```

------------------------------------------------------------------------

# Part II. ANOVA

## Problems

### Problem 1 (2 points)

```{r}
library(ggpubr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(broom)
library(car)

data <- read.csv('C:\\Users\\HP\\Desktop\\PESU\\SEM - 5\\PES1UG20CS445\\DA\\1_b\\Scenario 1.csv')
anova_1 <- aov(No.of.items ~ POI, data = data)
summary(anova_1)
```

## Analysis:

1.. Scully can use one-way ANOVA Test, because we are analyzing a single independent variable.

2.. aov() is a function used by Scully for ANOVA in R programming

3..NULL HYPOTHESIS : It states that the means are the same.

ALTERNATE HYPOTHESIS : It states that the means of every group is different.

Output: Since the P-Vaue is greater than alpha(5%).We accept the NULL Hypothesis, which means that there is no difference between the means .

### Problem 2 (3 points)

```{r}
data=read.csv('C:\\Users\\HP\\Desktop\\PESU\\SEM - 5\\PES1UG20CS445\\DA\\1_b\\Scenario 2.csv')
anova_2=aov(No.of.items ~ POI*Priority,data=data)
summary(anova_2)
```

## Analysis:

1.. We choose two-way ANOVA because there are two independent variables 2.. aov() function is used in R programming to find the AVOVA 3.. NULL HYPOTHESIS: the means are the same ALTERNATE HYPOTHESIS : the means are different ouput : Since the P-value is very less then the alpha(5%) value, we reject the NULL Hypothesis. 4..Global mean is not considered

### Problem 3 (1 point)

```{r}
HSD=TukeyHSD(anova_2)
HSD
plot(HSD)
```

## Analysis:

This test compares a set of means for the given category of data we observe that the pairs-- high-critical,low-critical and med-critical are insignificant(i.e. have a p-value \< 0.05)
