---
title: "**UE20CS312 - Data Analytics**\n\n**Worksheet 3a - Basic Forecasting Techniques**"
subtitle: "PES University"
author: "'SUNDEEP A, Dept. of CSE - PES1UG20CS445'"
date: "2022-09-29"
output: pdf_document
urlcolor: blue
editor_options:
  markdown:
    wrap: 72
---

### Time Series Data and Basic Forecasting Techniques

Time Series data is any data that is collected at regular time
intervals, with changing observations at every time interval. Processing
time series data effectively can help gain meaningful insights into how
a quantity changes with time.

Forecasting a quantity into the future is an essential task, that
predicts future values at any particular time. Forecasts can be made
using various techniques like Exponential Smoothing to much more complex
techniques such as Recurrent Neural Networks. Let's try to process
time-series data and forecast values using basic techniques!

### Prerequisites

-   Revise the following concepts
    -   Components of Time-Series Data
    -   Single, Double and Triple Exponential Smoothing
    -   Regression (Refer to worksheets and slides from Unit 2)
    -   Croston's Forecasting
    -   Time-Series Accuracy Metrics

### Task

Let's imagine it is 2012 and you are in the market to buy an Orange
Ultrabook Laptop for college. But this laptop is rare to find and
expensive. You would want to put your Data Analytics skills to use, and
predict the best time to buy your laptop, such that you can get it for
the best price! You would also like to predict when the Orange Ultrabook
would be in stock and when it would have high demand.

An electronics store collected sales data for their store weekly, from
*February 2010* to *October 2012*, a period of 143 months. You have
gotten your hands on this, and will use it to predict how the prices
will change in the future.

The data for the following tasks can be downloaded from [this Github
Repository](https://github.com/Data-Analytics-UE20CS312/Unit-3-Worksheets/blob/master/3a%20-%20Basic%20Forecasting%20Techniques/sales.csv).

### Data Dictionary

    Date - Date on which data was collected (end of the week)
    Sales - Weekly sales of the store (in $)
    Holiday_Flag - Boolean Flag. 0 for normal week and 1 for holiday season
    Temperature - Average temperature during the week
    Fuel_Price - Average price during the week (in $/gallon)
    CPI - Consumer Price Index
    Unemployment - Average percentage of Unemployment in the city
    Laptop_Demand - Number of Orange Ultrabook laptops sold during the week

### Data Ingestion and Preprocessing

-   Read the file into a `data.frame` object

```{r csv}
df <- read.csv('sales.csv')
head(df)
```

-   Pick out the `Sales` column in the `data.frame`. Most of our
    time-series analysis will be done on this column.

```{r}
sales <- df$Sales
head(sales)
```

-   The `ts` function is used to create the `ts` object in R. Frequency
    is 52 as it is weekly data. The start is specified like
    `start= c(y, m, d)` as we are dealing with weekly data. If it was
    monthly data we can omit the `d` and for yearly data we can omit the
    `m` as well.(`c` is the combine function in R)

```{r}
sales_ts <- ts(sales, frequency = 52, start=c(2010, 2, 5))
head(sales_ts)
```

-   Visualize the Time-Series of `Sales` column

```{r}
plot.ts(sales_ts)
```

### Points

The problems in this worksheet are for a total of 10 points with each
problem having a different weightage.

-   *Problem 1*: 1 point
-   *Problem 2*: 3 points
-   *Problem 3*: 2 points
-   *Problem 4*: 2 point
-   *Problem 5*: 2 points

### Problem 1 (1 Point)

Decompose the `Sales` column into trend, seasonal and random components.
Plot these components as well (Hint: Look at the `decompose` function).

```{r}
decomp = decompose(sales_ts)
#seasonal Component
seasonal_sales = decomp$seasonal
plot(seasonal_sales)
#Trend Component
trend_sales = decomp$trend
plot(trend_sales)
#Random Component
random_sales = decomp$random
plot(random_sales)
```

### Problem 2 (3 Points)

-   Perform forecasts using Single, Double and Triple Exponential
    Smoothing.

-   Plot the forecasts of all three forecasts (different colours)
    against the true values. (Hint: use `lines`)

-   **Use only one function needed for all 3 forecasts**, only changing
    parameters to get each of the 3 models (Hint: Think about alternate
    names)

    ```{r}
    library(fpp2)

    #Performing Single Exponential Smoothing
    Single_exponential_smoothing <- HoltWinters(sales_ts,beta=FALSE, gamma=FALSE , start = 2010)
    #Performing  Double Exponential Smoothing
    Double_exponential_smoothing <- HoltWinters(sales_ts,gamma=FALSE , start = 2010)
    #Performing  Triple Exponential Smoothing
    Triple_exponential_smoothing <- HoltWinters(sales_ts)



    plot(sales_ts,main="Single Exponential Smoothing")
    lines(Single_exponential_smoothing$fitted[,1], col="blue")

    plot(sales_ts,main="Double Exponential Smoothing")
    lines(Double_exponential_smoothing$fitted[,1],col="red")

    plot(Triple_exponential_smoothing,main="Triple Exponential Smoothing")

    ```

### Problem 3 (2 Points)

-   Forecast `Sales` values by Regression using all other columns. Print
    summary of regression model.
-   Plot the predicted values against original as well. (Hint:
    Regression model predictions will not be a Time Series, so need to
    get both to common index/x-axis)
-   (Hint: Will not work unless one column is dropped/transformed before
    including it in the regression. Use the `lm` function to get linear
    model)

Note: This is Multiple Linear Regression, that is, using all the columns
for regression

```{r}
lm_pred <- lm(formula=Sales ~ .-X-Date,data=df)
summary(lm_pred)
#Plot the predicted values against original
library(ggplot2)
ggplot(df, aes(x=predict(lm_pred), y= Sales)) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  labs(x='Predicted Values', y='Actual Values', title='Predicted vs. Actual Values')
```

### Problem 4 (2 Points)

Plot the `Laptop_Demand` column as a time series. Identify the
forecasting required for this type of Time-series, and forecast the
values for all 143 weeks (Hint: Look at functions in the `forecast`
package)

```{r}
library("forecast")
#Plotting the Laptop_Demand as a time Series
lpdmd_ts <- ts(df$Laptop_Demand, frequency = 52, start=c(2010, 2, 5))
plot.ts(lpdmd_ts)

#Forecasting required for this type of time-series is Holt-Winter Forecasting.
lpdmd_forecast <- forecast(lpdmd_ts, h=24, level=95)
plot(lpdmd_forecast,main='Holt-Winter Forecasts')
lines(lpdmd_forecast$fitted, lty=2, col="purple")
```

### Problem 5 (2 Points)

Evaluate the accuracy of all 3 Exponential Smoothing models (from
Problem 2) and Regression models using the MAPE and RMSE metrics.
Comment on which is the best Exponential Smoothing method, and if
Regression is better than Exponential Smoothing? Provide a reasoning for
why the best model is better suited for the `Sales` data (Bonus Point:
reasoning for why the 2 other models perform similarly)

```{r}
library("Metrics")
#For regression Model
sprintf("MAPE value for regression model is: %f",mape(sales_ts,lm_pred$fitted))
sprintf("RMSE value for Regression model is: %f",rmse(sales_ts,lm_pred$fitted))

#For Single Exponential Smoothing Model
sprintf("MAPE value for Single Exponential smoothing model is: %f",mape(sales_ts,Single_exponential_smoothing$fitted))
sprintf("RMSE value for Single Exponential smoothing model is: %f",rmse(sales_ts,Single_exponential_smoothing$fitted))

#For Double Exponential Smoothing Model
sprintf("MAPE value for Double Exponential smoothing model is: %f",mape(sales_ts,Double_exponential_smoothing$fitted))
sprintf("RMSE value for Double Exponential smoothing model is: %f",rmse(sales_ts,Double_exponential_smoothing$fitted))

#For Triple Exponential Smoothing Model
sprintf("MAPE value for Triple Exponential smoothing model is: %f",mape(sales_ts,Triple_exponential_smoothing$fitted))
sprintf("RMSE value for Triple Exponential smoothing model is: %f",rmse(sales_ts,Triple_exponential_smoothing$fitted))
```

**Analysis:**

Since, both MAPE and RMSE are error metrics, lower the error metrics
(i.e. MAPE and
RMSE), better is the prediction of the model.

Since Regression Model has the **lowest MAPE and RMSE** among all the
models. So we can say that Regression model is the Best Model.

In the Exponential Models, Single Regression Model had the Lowest MAPE
and RMSE values . So, we can say that Single Exponential Smoothing Model
is the Best model among the Exponential Smoothing Models.
